# Assignment 1

## Load data

```{r}
data <- read.csv("activity.csv", header = TRUE)
daily_steps <- aggregate(steps ~ date, data = data, sum)
```

## What is the mean total number of steps taken per day?

### Sum number of steps, create histogram and report median and mean.

```{r}
hist(daily_steps$steps, main = paste("Total Number of Steps per Day"), col = "gray", xlab = "Number of steps")
```

```{r}
daily_steps_mean <- mean(daily_steps$steps)
daily_steps_median <- median(daily_steps$steps)
```

The mean is `r daily_steps_mean`.
The median is `r daily_steps_median`.

# What is the average daily activity pattern?

## Time series plot and find interval with most average steps

```{r}
interval_steps <- aggregate(steps ~ interval, data = data, mean)
plot(interval_steps$interval, interval_steps$steps, type = "l", xlab = "5-minute interval", ylab = "Average steps taken", main = paste("Average Daily Activity Pattern"))

max_interval <- interval_steps[which.max(interval_steps$steps),1]
```

The mean is `r max_interval`.

# Imputing Missing Values

## Calculating missing values

```{r}
missing <- sum(!complete.cases(data))

# Loop thru all the rows of activity, find the one with NA for steps.
# For each identify the interval for that row
# Then identify the avg steps for that interval in avg_steps_per_interval
# Substitute the NA value with that value

for (i in 1:nrow(data)) {
    if(is.na(data$steps[i])) {
        val <- interval_steps$steps[which(interval_steps$interval == data$interval[i])]
        data$steps[i] <- val 
    }
}

daily_steps_impute <- aggregate(steps ~ date, data = data , sum)
```
## Create histogram

```{r}
hist(daily_steps_impute$steps, main = "Histogram of total number of steps per day (IMPUTED)", xlab = "Steps per day", col = "gray")
```

## Median and total number of steps
```{r}
dailystepsimpute_mean <- round(mean(daily_steps_impute$steps))
dailystepsimpute_median <- median(daily_steps_impute$steps)
```
The mean is equal to `r dailystepsimpute_mean` and the median is equal to `r dailystepsimpute_median`. It is important to note that the mean and median have not changed with the imputed numbers.

# Are there diferences in activity patterns between weekdays and weekends?

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
data$dow = as.factor(ifelse(is.element(weekdays(as.Date(data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```